<!--  common abbreviations: 
This is the snippet for summary of cluster profiler results
January Weiner

    cps, cluster profiler summary
    cp, cluster profiler 
-->


## Cluster profiler summary overview by database


### Overview




```{r cps_configuration}
require(tidyverse)
require(pander)
require(tmod)
require(orthomapper)
require(glue)

if(packageVersion("tmod") < "0.43") 
  stop("tmod version required: at least 0.43. Please install newer tmod from https://github.com/january3/tmod")

# load tmod functions
r.path <- "{{R_COMMON}}"
script_f <- file.path(r.path, "..", "..", "external_scripts", "tmod_functions.R")
source(script_f)

## config is defined in the main report template
if(is.null(config$contrasts$contrast_list)) stop("No contrasts defined")

contrasts <- sapply(config$contrasts$contrast_list, `[[`, "ID")
```


```{r cps_load_data}
## load cluster profiler data
cp_data <- lapply(contrasts, function(.c) {
  fn <- subset(file_tab, contrast == .c & step == "cluster_profiler" & extension == "rds")$filename
  unlist(readRDS(fn), recursive=FALSE)
})
names(cp_data) <- contrasts

cps_db_names <- unique(unlist(lapply(cp_data, names)))

## cp_data will now hold one element per each unique database name (e.g.
## MSigdb.H) and each of these elements has the results of enrichment
## analysis for each contrast for which it was run
cp_data <- lapply(cps_db_names, function(.n) {
  ret <- lapply(cp_data, `[[`, .n)
  ret[!(sapply(ret, is.null))]
})
names(cp_data) <- cps_db_names
```




```{r cps_deseq2_results}
message("Loading deseq2 results")
tmp <- subset(file_tab, step=="contrasts_full" & extension == "rds")
tmp2 <- lapply(tmp$filename, readRDS)
names(tmp2) <- tmp$contrast

.gl <- rownames(tmp2[[1]])
tmp2 <- lapply(tmp2, function(x) x[ .gl, ])

lfcs <- sapply(tmp2, `[[`, "log2FoldChange")
pvals <- sapply(tmp2, `[[`, "padj")
pvals[ is.na(pvals) ] <- 1
lfcs[ is.na(lfcs) ] <- 0
rownames(lfcs) <- .gl
rownames(pvals) <- .gl
```

**Table.** Overview of the databases for which gene set enrichment using
`cluster_profiler` was performed.

```{r cps_dbs_overview_table}

```


```{r cps_db_output_templates}
## Note: the `\n` character precedes the ```, because otherwise 
## knitr gets confused!
template_summary <- function(res, db_n) {
"

### `r db_n`

"
}

template_figure <- function(res, resS, db_n, effect.col) {

if(effect.col == "NES") { ret <-
"
**Fig.** Panel plot showing results for the database `r db_n`. Effect size
is the normalized enrichment score (NES). Blue color indicates negative
enrichment score, red color indicates positive NES. Size of the dots
corresponds to the magnitude of NES as shown in the legend. Color intensity
indicates p-value.

\n```{r, fig.width=9,fig.height=ceiling(4 + nrow(resS) / 2)}
labs <- tmod_labels(resS$ID, resS$Title, max.length=25, nlines=2)

tmodPanelPlot(resS, min.e=-3, max.e=3, filter.rows.auc=-Inf, pie.style='dotsymm', pval.thr=.05, row.labels=labs)
\n```
"
} else { ret <-
"
**Fig.** Panel plot showing results for the database `r db_n`. Effect size
is the relative enrichment score (E) defined as (b/n)/(B/N), where 
b is the number of significant genes in the given gene set, n is total
number of genes in the given gene set, B is the total number of significant
genes and N is the total number of genes.  Size of the dots corresponds to
the magnitude of E as shown in the legend. Color intensity indicates
p-value.

\n```{r, fig.width=9,fig.height=ceiling(4 + nrow(resS) / 2)}
labs <- tmod_labels(resS$ID, resS$Title, max.length=35, nlines=2)

ln <- names(labs)
labs <- sprintf('%s (%s)', labs, names(labs))
names(labs) <- ln
tmodPanelPlot(resS, filter.rows.auc=-Inf, pval.thr=.05, row.labels=labs, max.e=NULL, min.e=1)
\n```
"
}

return(ret)
}

```

```{r cps_results,results="asis"}
## Here, we knit the template for each database

for(db_n in names(cp_data)) {
  res <- lapply(cp_data[[db_n]], function(x) { x <- x@result ; x$Title <- x$Description ; x })
  message(glue("generating plots for {db_n}"))
 
  ## results with the "sort type" level flattened
  ## so instead of res$contrasts_name$pval$... we have res$contrasts_name.pval$...
  summary_t <- template_summary(res, db_n)
  foo <- knitr::knit_child(text=summary_t, quiet=TRUE)
  cat(foo)
 
  if("NES" %in% colnames(res[[1]])) {
    effect.col <- "NES"
  } else {
    effect.col <- "E"
    res <- lapply(res, function(x) {
      b <- as.numeric(gsub("/.*", "", x$GeneRatio))
      n <- as.numeric(gsub(".*/", "", x$GeneRatio))
      B <- as.numeric(gsub("/.*", "", x$BgRatio))
      N <- as.numeric(gsub(".*/", "", x$BgRatio))
      x$E <- ((b/n)/(B/N))
      x
    })
  }
    
  res_selection <- tmod_mod_sel(res, qval.thr=0.05, auc.thr=0, effect.col=effect.col, pval.col='p.adjust')
  res <- lapply(res, function(x) x[ x$ID %in% res_selection$sel_ids, , drop=FALSE ])
  resS <- tmodSummary(res, effect.col=effect.col, pval.col='p.adjust')

  if(!is.null(res_selection$note)) {
    cat(paste("\n\nNo figure produced because there were", res_selection$note, "enrichment results.\n\n"))
  } else {
    figure_t <- template_figure(res, resS, db_n, effect.col)
    foo <- knitr::knit_child(text=figure_t, quiet=TRUE)
    cat(foo)
  }
}
```






