



### Discordance / Concordance plots

```{r disco_template,fig.width=6,fig.height=6}

template <- '
\n```{r, results="asis"}
    cat(sprintf("\n###### %s\n", disco_title))

    c1 <- disco_contrasts[1]
    c2 <- disco_contrasts[2]
    cc <- merge(as(contrasts[[c1]], "data.frame"), as(contrasts[[c2]], "data.frame"), by=0)
    cc <- cc %>% replace_na(list(
      log2FoldChange.x=0,
      log2FoldChange.y=0,
      pvalue.x=1,
      pvalue.y=1))

    cc$disco <- with(cc, log2FoldChange.x * log2FoldChange.y * (-log10(pvalue.x) -log10(pvalue.y)))
    cc$disco[ is.na(cc$disco) ] <- 0
    cc$col <- disco.map(cc$disco)
    cc <- cc %>% arrange(abs(cc$disco))

    #plot(cc[["log2FoldChange.x"]], cc[["log2FoldChange.y"]], col=cc$col, pch=19)
    ggplot(cc, aes(x=log2FoldChange.x, y=log2FoldChange.y, col=col)) + 
    geom_point() +
    scale_color_manual(values=
      set_names(
        unique(cc$col), unique(cc$col)
      )
    ) + theme(legend.position="none")+
    xlab(c1) +
    ylab(c2) +
    ggtitle(
      sprintf("Correlation between log2FC for %s and %s\nr=%.2f rho=%.2f",
        c1, c2,
        cor(cc$log2FoldChange.x, cc$log2FoldChange.y, use="p"),
        cor(cc$log2FoldChange.x, cc$log2FoldChange.y, method="s", use="p"))) +
    geom_hline(aes(yintercept=0), color="grey") +
    geom_vline(aes(xintercept=0), color="grey") +
    geom_abline(aes(slope=1, intercept=0), color="grey")


\n```

'




```


```{r disco}
require(rlang)
require(purrr)
require(dplyr)
require(ggplot2)
require(tidyr)

disco <- config$report$snippet_parameters$Disco

cntr_ids <- map_chr(config$contrasts$contrast_list, ~ .x[["ID"]])
cntr_nms <- map_chr(config$contrasts$contrast_list, ~ .x[["title"]])
names(cntr_ids) <- cntr_nms

contrasts <- file_tab %>% filter(step == "contrast" & extension == "rds") 
contrasts <- contrasts %>% pull(filename) %>% set_names(contrasts[["contrast"]]) %>% map(readRDS)

disco_n <- length(disco)


disco.map <- function(x, lower=-10, upper=10, int=255) {
  x[ is.na(x) ] <- 0
  x[ x > upper ] <- upper
  x[ x < lower ] <- lower

  pal <- colorRampPalette(c("blue", "grey", "red"))(int)
  pal <- paste0(pal, "66")

  int <- findInterval(x, seq(lower, upper, length.out=int + 1), all.inside=TRUE)
  pal[int]
}

```

```{r disco_plots, results="asis"}
cat("\n##### {.tabset}\n")

for(d in disco) {
  disco_contrasts <- cntr_ids[ d[["contrasts"]] ]
  disco_title <- d[[ "title" ]]

  out <- knitr::knit_child(text=template, quiet=TRUE)
  cat(out)



}

cat("\n##### {-}\n")
```




