#REQUIRE {{tmod_pca-rds}}

<!--
This is the snippet showing the results of tmodPCA
January Weiner
-->

## tmodPCA analysis


### Overview


```{r tmod_pca_configuration}
## load the necessary libraries and objects required to run this snippet
require(tidyverse)
require(pander)
require(tmod)
require(orthomapper)
require(glue)

## get the file names from file_tab
#tmod.res.f <- subset(file_tab, step=="tmod_summary" & extension == "rds")$filename
tmod.dbs.f <- subset(file_tab, step=="tmod_dbs" & extension == "rds")$filename
map.f      <- subset(file_tab, step == "tmod_dbs" & extension == "mapping.rds")$filename
annot.f    <- subset(file_tab, step == "annotation" & extension == "rds")$filename
tmod.pca.f <- subset(file_tab, step == "tmod_pca" & extension == "rds")$filename
tmod.pca.obj.f <- subset(file_tab, step == "tmod_pca" & extension == "pca.rds")$filename

#tmod.res  <- readRDS(tmod.res.f)
tmod.dbs  <- readRDS(tmod.dbs.f)
conf.snip <- config$report$snippet_parameters$tmod
annot     <- readRDS(annot.f)
pca       <- readRDS(tmod.pca.obj.f)
tmod.pca  <- readRDS(tmod.pca.f)

tmod.mapping <- readRDS(map.f)

# load tmod functions
r.path <- r_common
script_f <- file.path(r.path, "..", "..", "external_scripts", "tmod_functions.R")
source(script_f)
```

```{r tmod_pca_template}

## variables required:
## db_n name of the database
## res the enrichment results
## params params for the plot
template <- "

Figure: database `r db_n`

\n```{r}
N <- floor(length(res)/2)


\n```

\n```{r,fig.width=12,fig.height=12}
tmodpca_obj <- list()
tmodpca_obj$params <- params
class(tmodpca_obj) <- c('tmodPCA', class(tmodpca_obj))

N <- 1
for(i in 1:N) {
  cc <- c(i*2 - 1, i*2)
  .res <- res[paste0('PC.', cc)]
  .restmp <- unlist(.res, recursive=FALSE)
  ids <- tmod_mod_sel(.restmp, min.n=15, max.n=30)$sel_ids
  .res <- lapply(.res, function(xx) lapply(xx, function(x) {
    ret <- x[ x$ID %in% ids, , drop=FALSE ]
    ret$Title <- tmod_labels(ret$ID, ret$Title)
    ret
    }))
  tmodpca_obj$enrichments <- .res
  tmodpca_obj$params$components <- cc
  #plot(tmodpca_obj)
  plot(1:10)
}
\n```

"



```

```{r tmod_pca_figures,fig.width=12,fig.height=12}

params <- list(
  pca=pca,
  filter=TRUE,
  simplify=TRUE,
  legend=FALSE,
  plot=TRUE,
  mode="leftbottom",
  plotfunc=pcaplot)


for(db_n in names(tmod.pca)) {
  res <- tmod.pca[[db_n]]

  foo <- knitr::knit_child(text=template, quiet=TRUE)

}

```


