<!--
This is the snippet for summary of tmod results.
January Weiner
-->

## Gene set enrichment analysis with tmod


### Overview


```{r tmod_configuration}
## load the necessary libraries and objects required to run this snippet
require(tidyverse)
require(pander)
require(tmod)
require(orthomapper)
require(glue)
require(purrr)

## get the file names from file_tab
#tmod.res.f <- subset(file_tab, step=="tmod_summary" & extension == "rds")$filename
tmod.dbs.f <- subset(file_tab, step=="tmod_dbs" & extension == "rds")$filename
map.f      <- subset(file_tab, step == "tmod_dbs" & extension == "mapping.rds")$filename
annot.f    <- subset(file_tab, step == "annotation" & extension == "rds")$filename

#tmod.res  <- readRDS(tmod.res.f)
tmod.dbs  <- readRDS(tmod.dbs.f)
conf.snip <- config$report$snippet_parameters$tmod
annot     <- readRDS(annot.f)

tmod.mapping <- readRDS(map.f)

# load tmod functions
r.path <- r_common
script_f <- file.path(r.path, "..", "..", "external_scripts", "tmod_functions.R")
source(script_f)
```


```{r tmod_summary}
## create tmod_res with all the results
contr <- sapply(config$contrasts$contrast_list, `[[`, "title")
names(contr) <- sapply(config$contrasts$contrast_list, `[[`, "ID")

res.f <- subset(file_tab, step == "tmod" & extension == "rds")$filename
cnames <- subset(file_tab, step == "tmod" & extension == "rds")$contrast

all_res <- lapply(res.f, readRDS)
names(all_res) <- cnames

contr <- contr[ names(contr) %in% cnames ]

## put the contrasts in the same order they are in the config file
all_res <- all_res[names(contr)]

res <- lapply(names(tmod.dbs), function(n) lapply(all_res, `[[`, n))
names(res) <- names(tmod.dbs)

tmod.res <- list(results=res, contrasts_names=contr)
rm(res)



```



```{r tmod_genelists}
message(glue("Loading tmod gene lists"))

tmp <- subset(file_tab, step=="tmod" & extension == "gl.rds")
genelists <- lapply(tmp$filename, readRDS)
names(genelists) <- tmp$contrast

dbs <- unique(unlist(lapply(genelists, names)))

## genelists holds the ordered lists of genes corresponding to contrasts and 
## enrichment methods. Each element corresponds to the database and is itself a list 
## with one element per contrast.
genelists <- lapply(dbs, function(x) {
  x <- lapply(genelists, `[[`, x)
  x[names(contr)]
})
names(genelists) <- dbs
```


```{r deseq2_results}
## Load the deseq2 results. We need them to visualize the data.
message("Loading deseq2 results")
tmp <- subset(file_tab, step=="contrasts_full" & extension == "rds")
tmp2 <- lapply(tmp$filename, readRDS)
names(tmp2) <- tmp$contrast

.gl <- rownames(tmp2[[1]])
tmp2 <- lapply(tmp2, function(x) x[ .gl, ])

lfcs <- sapply(tmp2, `[[`, "log2FoldChange")
pvals <- sapply(tmp2, `[[`, "padj")
pvals[ is.na(pvals) ] <- 1
lfcs[ is.na(lfcs) ] <- 0
rownames(lfcs)  <- .gl
rownames(pvals) <- .gl
```

**Table.** Overview of the databases for which gene set enrichment using
tmod was performed.

```{r tmod_dbs_overview_table}
dbstat <- function(db) {
  n <- nrow(db$dbobj$MODULES)
  data.frame(ID=db$name, Name=db$title, Description=db$description, TaxonID=db$taxonID, N=n)
}

tab <- lapply(tmod.dbs, dbstat)
tab <- Reduce(rbind, tab)
pander(tab, row.names=FALSE, split.tables=Inf)
```


```{r tmod_db_output_templates}
## we create templates as functions only to indicate which variables 
## the templates need
## Note: the `\n` character precedes the ```, because otherwise 
## knitr gets confused!

template_summary <- function(db_n, db, resU) {
"

## tmod enrichment analysis results for database `r db$title`.

### Summary

**Database ID:** `r db$name`.

**Description:** `r db$description`.

**Tab. Summary of the results.** Numbers show the number of enrichments
significant at a given threshold for the given contrast and test type.

\n```{r, results='asis'}
pval.thr <- c(.05, .01, 1e-3, 1e-5)
tab <- imap_dfr(resU, ~ {
  .res   <- .x
  .cname <- .y

  map_int(pval.thr, ~ sum(.res$adj.P.Val < .x, na.rm=TRUE))
})

tab <- data.frame(names(resU), t(tab))
colnames(tab) <- c('Contrast', pval.thr)

pander(tab, row.names=FALSE, split.tables=Inf)
\n```

"
}

template_figure <- function(db_n, db, res, resU, lfcs, pvals, tmod.mapping, res_selection) {
"

### Figure

\n```{r}
## here, we run tmodDecideTests to figure out which genes belonging to a module go up, which go down

## translate the names of the DESeq2 results to tmod database IDs
gl <- rownames(lfcs)
tmod.mapname <- tmod.mapping$dbs[db_n]
gl.mapped <- tmod.mapping$maps[[tmod.mapname]][gl]

## run tmodDecideTests to get numbers of significantly regulated genes 
pie.main <- tmodDecideTests(gl.mapped, lfcs, pvals, mset=db$dbobj[ res_selection$sel_ids ] )

## Since we can have multiple tests per database, but only one 'pie' per
## contrast, we need to recreate the structure of tmod.res

pies <- imap(res, ~ {
  res.ctr <- .x
  ctr_name <- .y
 
  map(res.ctr, ~ {
    pie.main[[ctr_name]]
  })
})

## object analogous to resU, it will have the same names
piesU <- unlist(pies, recursive=FALSE)
\n```

**Fig.** Panel plot showing results for the database `r db$name`.

\n```{r}
resU2 <- lapply(resU, function(x) x[ x$ID %in% res_selection$sel_ids,, drop=FALSE ])
resU2S <- tmodSummary(resU2)
rowlabs <- tmod_labels(resU2S$ID, resU2S$Title, max.length=35, nlines=2)
#message(sprintf('Num of results to show: %d', nrow(resU2S)))
#pander(resU2S)
\n```


\n```{r, fig.width=12,fig.height=4 + ceiling(nrow(resU2S)/2)}
tmodPanelPlot(resU2, filter.rows.pval=conf.snip$fig_qval_max, 
                     filter.rows.auc=conf.snip$fig_auc_min, 
										 pie=piesU, grid='b', row.labels=rowlabs, text.cex=0.7)
\n```
"
}

template_evidence_plots <- function(db_n, db, ids, gl) {
"

\n```{r}
n <- length(ids)
## collapse the top two levels of genelists
glU <- unlist(gl, recursive=FALSE)
\n```

### Evidence plots

Figures below show the evidence plots for the top `r n` gene sets. Each row
corresponds to one gene set. Each column corresponds to one enrichment test
(contrast + ordering). Each evidence plot shows the existing evidence for the
enrichment of the given gene set in the given contrast. The curve shows
the Receiver Operator Characteristic (ROC) curve for a given gene set. The rug
below the figure represents the ordered list of genes. Genes belonging to a 
given gene set are highlighted. Colors indicate whether the genes are
positively or negatively regulated (red or blue, respectively), while color
brightness indicates whether genes are significantly regulated (at q < 0.05).



\n```{r fig.width=4*length(glU), fig.height=4*n}

par(mfrow=c(n, length(glU)))

## color the genes
## names(gl) are the contrast names

gids <- rownames(lfcs)
tmod.mapname <- tmod.mapping$dbs[db_n]
gids.mapped <- tmod.mapping$maps[[tmod.mapname]][gids]

gene_colors <- lapply(names(gl), function(.gl) {
	p <- pvals[,.gl]
	l <- lfcs[,.gl]

	cols <- ifelse(l < 0, 
		ifelse(p < .05, 'blue', '#660000'),
		ifelse(p < .05, 'red', '#000066'))
	names(cols) <- gids.mapped
	cols
})
names(gene_colors) <- names(gl)

mlabs <- tmod_labels(ids, db$dbobj$MODULES[ids, 'Title' ], max.length=35, nlines=2)

gsymbols <- NULL
if('SYMBOL' %in% colnames(annot)) {
	gsymbols <- annot$SYMBOL
	names(gsymbols) <- annot$PrimaryID
	gsymbols[is.na(gsymbols)] <- ''
}
	

for(id in ids) {
  for(.c in names(gl)) {
		for(.s in names(gl[[.c]])) {
			glabs <- gl[[.c]][[.s]] # gene labels
			glabs <- glabs[ glabs %in% db$dbobj$MODULES2GENES[[id]] ]
			if(!is.null(gsymbols)) {
				gene_labs <- gsymbols[ names(glabs) ]
			} else {
				gene_labs <- glabs
			}
			names(gene_labs) <- glabs

    	evidencePlot(gl[[.c]][[.s]], 
				m=id, 
				mset=db$dbobj, 
				gene.colors=gene_colors[[.c]],
				gene.labels=gene_labs,
				main=sprintf('%s / %s\n%s', contr[.c], .s, mlabs[id]))
		}
  }
}


\n```

"
}

```

```{r tmod_results,results="asis"}
## Here, we knit the template for each database

for(db_n in names(tmod.res$results)) {
  res  <- tmod.res$results[[db_n]]
  db <- tmod.dbs[[ db_n ]]
  message(glue("generating plots for {db_n}"))

  ## results with the "sort type" level flattened
  ## so instead of res$contrasts_name$pval$... we have res$contrasts_name.pval$...
  resU <- unlist(res, recursive=FALSE)
  resU <- resU[ !sapply(resU, is.null) ]

  if(length(resU) < 1L) {
    cat("\n\nNo figure produced because there were no enrichment results.\n\n")
  } else {
    summary_t <- template_summary(db_n, db, resU)
    foo <- knitr::knit_child(text=summary_t, quiet=TRUE)
    cat(foo)

    # tmod_mod_sel <- function(res, qval.thr=.01, auc.thr=.65, max.n=25, min.n=max.n, res_sum=NULL) {
    res_selection <- tmod_mod_sel(resU, qval.thr=conf.snip$fig_qval_max, 
                                       auc.thr=conf.snip$fig_auc_min,
                                       max.n=conf.snip$fig_n_max,
                                       min.n=conf.snip$fig_n_min)
    if(!is.null(res_selection$note)) {
      cat(paste("\n\nNo figure produced because there were", res_selection$note, ".\n\n"))
    } else {

      if(length(res_selection$sel_ids) < 2) {
        cat(paste("\n\nNo panel plot produced because there was only 1 module to show.\n\n"))
      } else {
        figure_t <- template_figure(db_n, db, res, resU, lfcs, pvals, tmod.mapping, res_selection)
        foo <- knitr::knit_child(text=figure_t, quiet=TRUE)
        cat(foo)
      }

      n <- conf.snip$n_evid
      .sel <- tmod_mod_sel(resU, qval.thr=conf.snip$fig_qval_max, auc.thr=conf.snip$fig_auc_min, max.n=n, min.n=n)
      ids <- .sel$sel_ids
      gl <- genelists[[db_n]]
      evidence_plot_t <- template_evidence_plots(db_n, db, ids, gl)
      foo <- knitr::knit_child(text=evidence_plot_t, quiet=TRUE)
      cat(foo)
    }
  }
}
```






