## Principal component analysis

The principal component analysis plot shown below was generated using the most varying `r config$report$snippet_parameters$Normalisation_QC$n_most_varying` genes across all samples.

The expression values are obtained by the "`r config$normalization$normalized_expression`" method, where the normalisation doesn't take into account the experimental design.

In presence of strong biological signal, the samples should cluster with the biological condition.
When samples are clustered according to other effects (for example patient, or technical batch), great care must be used when interpreting the results, as the other effects will considerably reduce the ability to extract meaningful biological information.


<!-- Compute PCA using most varying genes or all if config$report$snippet_parameters$Normalisation_QC$n_most_varying == 0 -->
```{r normalisationQCPCA_calc}
mtx <- t(assay(RLD_BLIND))
vars <- sort(apply(mtx, 2, var), decreasing=TRUE)
sel <- names(vars)[vars > 1e-26]
if (config$report$snippet_parameters$Normalisation_QC$n_most_varying > 0)
	sel <- sel[1:min(length(sel), config$report$snippet_parameters$Normalisation_QC$n_most_varying)]
pca <- prcomp(mtx[ , sel], scale.=TRUE)
```

<!-- Prepare for the plotting:
     - Extract the covariates to display in the PCA plot, as per config$report$snippet_parameters$Normalisation_QC$annotation_columns,
	   and merge with the principal components. 
	   The points' colours will be assigned to the first covariate, and the points' shapes to the second. Extra covariates will be ignored.
	   By default, config$report$snippet_parameters$Normalisation_QC$annotation_columns should contain only "group".
     - Please note that the row names of the data frame colData(RLD_BLIND) must contain the sample labels, identical to the column names
	   of the counts and expression matrices.
	 - Because of plot_ly internals, the number of distinct groupings (n_group) must be computed. It is the product between the number of
	   levels in the colour & shape covariates.
-->

```{r nomalisationQCPCA_prepare}
library(dplyr)
library(tibble)
library(purrr)

numplots <- floor(ncol(pca$x)/2)
numplots <- min(numplots, 6)

covariate_data <- as.data.frame(colData(RLD_BLIND))

cov_default <- config$report$snippet_parameters$Normalisation_QC$annotation_columns
threeD <- config$report$snippet_parameters$Normalisation_QC$threeD %||% FALSE

df <- data.frame(
	label=rownames(covariate_data),
	pca$x[rownames(covariate_data), paste("PC", 1:(2*numplots), sep="")],
	stringsAsFactors=FALSE
)

## selecting covariates to use: not all unique and also not only one value
sel <- apply(covariate_data, 2, function(x) {
  .u <- unique(x)
  length(.u) > 1 && length(.u) < nrow(covariate_data)
})

covariates <- union(cov_default, colnames(covariate_data)[sel])
covariate_data <- covariate_data[ , covariates, drop=FALSE ]

## we are sure that the order is the same
df <- cbind(covariate_data, df)

## which covariates are numerical?
cov_numeric <- covariates[ map_lgl(covariates, ~ is.numeric(df[[.]])) ]

## make sure the covariates are factors
df <- df %>% mutate_at(setdiff(covariates, cov_numeric), factor)


## make the PCA data more terse
df <- df %>% mutate_at(vars(starts_with("PC")), signif, digits=4)

## auto select default covariates if cov_default is missing
if(length(cov_default) < 1) {
  if(length(covariates) > 0) {
    cov_default <- covariates[1]
  }
}

if(length(cov_default) < 2) {
  if(length(covariates) > 1) {
    cov_default <- covariates[2]
  }
}
```

<!-- Plotting. As an example, the code below produces up to 6 projections (i.e. up to the 12th principal component).
Two distinct output are produced here:
- A pdf file, which contains up to 6 plots, components 1 vs 2, 3 vs 4, ...
  The filename is PLOTS_DIR/PCA.pdf, there is one page per projection, and the size is the default size of the pdf function.
- One interactive plotly figure embedded in the report. The user can choose which projection she wants to see,
  and the gene label is shown when the pointer hovers over the point.
-->

```{r normalisationQCPCA_plot_PDF, fig.height=7, fig.width=7}
## first, the PDF plot.
pdf_plot_f <- function(df, x, y, cov_default) {
  require(ggplot2)

  color <- cov_default[1] %||% NULL
  shape <- cov_default[2] %||% NULL

	ggplot(df, ggplot2::aes_string(x=x, y=y, color=color, shape=shape)) +
		geom_point() +
		labs(color=color, shape=shape) +
		coord_fixed()
}

fn <- file.path(PLOTS_DIR, "PCA.pdf")
pdf(fn)

for (i in 1:numplots) {
  print(pdf_plot_f(df, paste("PC", 2*i-1, sep=""), paste("PC", 2*i, sep=""), cov_default))
}
dev.off()
```

```{r normalisationQCPCA_plot_plotly, fig.height=7, fig.width=10}
library(plotly)
library(RColorBrewer)

plotly_ids <- paste0("ID", 1:nrow(df))

## First, assign palettes to factor covariates
## We need to set up coloring manually, because plotly is stupid
## use palettes from RColorBrewer, preferring the colorblind ones
palettes <- brewer.pal.info %>% rownames_to_column("palette") %>% filter(category == "qual") %>% arrange(-colorblind)
palette_names <- palettes[["palette"]]

covariates_fac <- covariates[ sapply(covariates, function(cov) is.factor(df[[cov]])) ]
if(nrow(palettes) < length(covariates_fac)) {
  palette_names <- rep(palette_names, ceiling(length(covariates_fac)/nrow(palettes)))
}

covariates_pal_names <- rlang::set_names(palette_names[1:length(covariates_fac)], covariates_fac)

## get colors for each of the factor covariates
covariates_col <- lapply(rlang::set_names(covariates_fac), function(cov) {
  pal.name <- covariates_pal_names[cov]
  n.col <- length(levels(df[[cov]]))
  max.col <- subset(palettes, palette == pal.name)$maxcolors[1]
  .pal <- brewer.pal(n=min(max.col, max(3, n.col)), name=pal.name)
  if(max.col < n.col) {
    warning(sprintf("covariate: %s more levels (%d) than colors (%d) in palette (%s)",
      cov, n.col, max.col, pal.name))
    .pal <- rep(.pal, ceiling(n.col / max.col))
  }
  names(.pal) <- levels(df[[cov]])

  rlang::set_names(.pal[ as.character(df[[cov]]) ], plotly_ids)
})

cont_pal <- colorRampPalette(c("#440154FF", "#21908CFF", "#FDE725FF"))(32)
covariates_col_cont <- lapply(rlang::set_names(setdiff(covariates, covariates_fac)), function(cov) {
  x <- df[[cov]]
  col_i <- seq(min(x, na.rm=TRUE), max(x, na.rm=TRUE), length.out=33)
  i <- findInterval(x, col_i, all.inside=TRUE)
  rlang::set_names(cont_pal[i], plotly_ids)
})

covariates_col <- c(covariates_col, covariates_col_cont)


## Same for symbols
symbols <- c("circle", "square", "diamond", "cross", "x", "triangle",
             "pentagon", "hexagram", "star", "diamond", "hourglass",
             "bowtie", "asterisk", "hash")

cov_symbols <- lapply(rlang::set_names(covariates), function(cov) {
  lev <- levels(df[[cov]])
  .s <- symbols
  if(length(lev) > length(.s)) {
    # expand symbol list if necessary. Moot, because if there are more
    # than three symbols it is pointless anyway
    .s <- rep(.s, ceiling(length(lev) / length(.s)))
  }

  .s <- .s[1:length(lev)]
  names(.s) <- lev
  return(.s)
})

plotly_ids <- paste0("ID", 1:nrow(df))

## create hovertext as a combination of covariates
hovertext <- apply(map_dfc(rlang::set_names(covariates), ~ sprintf("%s: %s", ., df[[.]])), 1, paste, collapse="\n")
hovertext <- paste(hovertext, sprintf("plotlyID: %s", plotly_ids), sep="\n")

## set up default colors and symbols

def_symbol <- def_symbols <- def_color <- def_colors <- NULL
if(length(cov_default) > 0) {
  def_color <- df[[ cov_default[1] ]]
  if(is.factor(def_color)) {
    levs <- levels(def_color)
    matches <- match(levs, df[[ cov_default[1] ]])
    def_colors <- covariates_col[[cov_default[1]]][ matches ]
    names(def_colors) <- levs
  } else {
    def_colors <- covariates_col[[cov_default[1]]]
  }
}



if(length(cov_default) > 1) {
  def_symbol <- df[[ cov_default[2] ]]
  def_symbols <- cov_symbols[[ cov_default[2] ]]
} 

if(threeD) {
  numplots <- floor(numplots * 2/3)
}

## create the initial plotly object. We need it to know how plotly cuts up the data into segments
plotly.args <- list(data=df, type="scatter", mode="markers",
             x=df$PC1, y=df$PC2,
             hovertext=hovertext,
             ids=plotly_ids,
             marker=list(size=10),
             visible=TRUE,
             color= def_color, colors= def_colors,
             symbol=def_symbol, symbols=def_symbols)

if(threeD) {
  plotly.args$z <- df$PC3
  plotly.args$type <- "scatter3d"
  plotly.args$marker <- list(size=5)
}

p <- do.call("plot_ly", plotly.args)


## if there are more than two components to be shown, we need to add the data
if(numplots > 1) {
  if(threeD) {
    for(i in 2:numplots) {
      pc1 <- paste0("PC", i * 3 - 2)
      pc2 <- paste0("PC", i * 3 - 1)
      pc3 <- paste0("PC", i * 3)
      p <- p %>% add_trace(data=df, type="scatter3d", mode="markers", x=df[[pc1]], y=df[[pc2]], z=df[[pc3]], visible=FALSE)
    }
  } else {
    for(i in 2:numplots) {
      pc1 <- paste0("PC", i * 2 - 1)
      pc2 <- paste0("PC", i * 2)
      p <- p %>% add_trace(data=df, type="scatter", mode="markers", x=df[[pc1]], y=df[[pc2]], visible=FALSE)
    }
  }
}


## here we build the plotly object and dissect it to figure out
## how plotly cut up the data. Note that plotly is efficient and that if
## factor levels overlap between factors (i.e. factors are colinear) then
## there will be fewer blocks than simply length(levels(f1)) * length(levels(f2))
## Also, we need the IDs in each block to be able to match the correct values
p.obj <- plotly::plotly_build(p)$x$data
nblocks <- length(p.obj) / numplots
p.obj <- p.obj[ 1:nblocks ]
p.ids <- map(p.obj, ~ .$ids)

## here we build the menu for updating symbol style
sym_update <- lapply(covariates_fac, function(cov) {
  sym_pal <- cov_symbols[[cov]]

  ## map levels to symbol
  syms <- lapply(p.ids, function(ids) {
    ord <- match(ids, paste0("ID", 1:nrow(df)))
    var <- df[[cov]][ ord ]
    sym_pal[ as.character(var) ]
  })

  list(
        method="restyle",
        label=sprintf("Symbol by %s", cov),
        args=list(list(
            marker.symbol=syms
            ))
      )
})

## set up update menus for colors
col_update <- lapply(covariates, function(cov) {
  cols <- lapply(p.ids, function(ids) {
    covariates_col[[cov]][ids]
  })
  list(
        method="restyle",
        label=sprintf("Color by %s", cov),
        args=list(list(
            marker.color=cols
            ))
      )
})

## prepare the buttons to change the coordinates
if(threeD) {
  pc_update <- lapply(1:numplots, function(i) {
    pc1 <- paste0("PC", i * 3 - 2)
    pc2 <- paste0("PC", i * 3 - 1)
    pc3 <- paste0("PC", i * 3)
    list(
      method="update",
      label=sprintf("PC%d/%d/%d", i * 3 - 2, i * 3 - 1, i * 3),
      args=list(
        list(visible=rep((1:numplots) == i, each=nblocks)),
        list(scene=list(
            xaxis=list(title=pc1),
            yaxis=list(title=pc2),
            zaxis=list(title=pc3)
          ))
      ))
  })
} else {
  pc_update <- lapply(1:numplots, function(i) {
    pc1 <- paste0("PC", i * 2 - 1)
    pc2 <- paste0("PC", i * 2)
    list(
      method="restyle",
      label=paste(pc1, "vs", pc2),
      args=list(
        list(visible=rep((1:numplots) == i, each=nblocks))
     ))
  })
}


menus <- list(
  list(
    type="buttons",
    direction="right",
    y=1.1,
    yanchor="top",
    x=0.1,
    xanchor="left",
    buttons=pc_update
  ),
  list(
    yanchor="bottom",
    y=0,
    x=-.1,
    xanchor="right",
    direction="up",
    buttons=col_update
  ),
  list(
    yanchor="bottom",
    y=.1,
    xanchor="right",
    x=-.1,
    direction="up",
    buttons=sym_update
  )
)

scene <- list(
  xaxis=list(title="PC1"),
  yaxis=list(title="PC2")
)

if(threeD) {
  scene$zaxis <- list(title="PC3")
}

p %>% layout(updatemenus=menus,
  scene=scene,
  legend=list(
    xanchor="right",
    x=-.1
  ))


```

[Download plot](`r fn`)
